---
title: Analysing the effect of land-use type on the proportion of scouts - a *jackknife approach*
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---

---

```{r preamble, include=FALSE}
devtools::load_all()
library(showtext)
showtext_auto()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic()
)

set_ggplot_theme <- function() {
  theme_set(
    theme_classic() +
      theme(
        text = element_text(family = "URWHelvetica", size = 18),
        legend.position = "none",
        # plot.title = element_text(hjust = 0.5),
        # axis.text.x = element_text(angle = 0,  hjust=1)
      )
  )
}

set_ggplot_theme()

knitr::opts_chunk$set(
  dev = "png",
  dpi = 300,
  collapse = TRUE,
  comment = "#>",
  fig.width = 12,
  fig.height = 8
)

library(cowplot)
library(tidyr)
library(dplyr)
library(ggfortify)
library(betareg)
library(plsRbeta)
library(lmtest)
library(kableExtra)
library(effects)
```

```{r get-data, include=FALSE, message = FALSE, error = FALSE, warning = FALSE}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("../Results/site_fit_results.Rda")

# filter for best model
sitedata <- sitedata %>%
  group_by(site) %>%
  slice(which.min(AIC))

# Load land use data taken (taken from Samuelson et al. 2021)
urban_lu <- read.csv(file = "../Data/urban-landuse.csv")
rural_lu <- read.csv(file = "../Data/agri-rural-landuse.csv")

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)

urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu) %>%
  filter(site != "ROT") # remove ROT as fit is bad

rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")


```


```{r, diff-p-sites}

urban_p <- urban_lu %>% select(site, p)
rural_p <- rural_lu %>% select(site, p)

urban_p$LU = "Urban"
rural_p$LU = "Agri-rural"

comb <- rbind(urban_p, rural_p) %>% mutate(p = 1-p)
comb_ar <- comb %>% filter(LU == "Agri-rural")
comb_u <- comb %>% filter(LU == "Urban")

plt <- ggplot(comb, aes(y = p, x = LU)) +
  geom_violin(fill="gray") +
  # geom_boxplot(width=0.05) +
  ylab("Waggle dance use") +
  xlab("Environment")

ggsave(
  plot = plt,
  filename = "../Results/figures/SI_figures/urban_rural_violin.png",
  dpi = 300
)

ggsave(
  plot = plt,
  filename = "../Results/figures/SI_figures/urban_rural_violin_nobox.svg",
  dpi = 300
)



res = chisq.test(comb_ar$p, comb_u$p)
res

```


```{r, convert_p}

ms_mr <- read.csv(file = "../Data/site_ms_mr.csv")

urban_lu <- urban_lu %>%
  # inner_join(ms_mr, by = "site") %>%
  # select(-ms, -mr, -p.y) %>%
  # rename(p = p_trips) %>%
  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
  mutate(p = 1-p)
  # select(site, p, 3:10)

rural_lu <- rural_lu %>%
  # inner_join(ms_mr, by = "site") %>%
  # select(-ms, -mr, -p.y) %>%
  # rename(p = p_trips) %>%
  # mutate(p = ifelse(p.x == 1, 1, p)) %>%
  mutate(p = 1-p)
  # select(site, p, woodland, 3:12)

```


# Jackknife analysis to investigate PLS sensitivity to outliers

```{r, jackknife_functions, message=FALSE, warning=FALSE, echo = FALSE}

ggplotRegression <- function(fit) {
  plt <- ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) +
    geom_point() +
    stat_smooth(method = "lm", col = "red") +
    labs(
      title = paste(
        "Adj R2 = ", signif(summary(fit)$adj.r.squared, 5),
        "Intercept =", signif(fit$coef[[1]], 5),
        " Slope =", signif(fit$coef[[2]], 5),
        " P =", signif(summary(fit)$coef[2, 4], 5)
      )
    )
  return(plt)
}

run_pls <- function(data) {
  model_pls <- plsRbeta(
    p_transform ~ .,
    data = data,
    nt = ncol(data) - 1,
    model = "pls-beta",
    verbose = FALSE
  )
  plsdata <- data.frame(
    prop.scouts = model_pls$dataY,
    pc1 = model_pls$tt[, 1]
  )
  plsbetamodel <- betareg(
    prop.scouts ~ pc1,
    plsdata
  )
  pls_loadings <- model_pls$pp[, 1]
  # get beta regression data
  plsdata <- data.frame(
    prop.scouts = model_pls$dataY,
    model_pls$tt
  )
  returnlist <- list(
    "loadings" = pls_loadings, "plsdata" = plsdata[, 1:2],
    "coef" = coef(plsbetamodel)[[2]]
  )
  return(returnlist)
}
```

## Urban jackknife

```{r, urban_pls, message = FALSE, error = FALSE, warning = FALSE, include = FALSE}

# set up data for pls
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]
sites <- urban_lu$site
urban_cuttoff <- sqrt(1 / (ncol(urban.pcadata) - 1))
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n
urban.pcadata$p_transform <- y

# create empty lists to store jackknifed results in
jkloadings <- list()
jkplsdata <- list()
jkcoefs <- list()

# jackknife the pls analysis
for (i in c(0, seq_len(nrow(urban.pcadata)))) {
  data <- urban.pcadata
  removed <- 0
  usesites <- "None"
  removed_site <- "None"
  if (i > 0) {
    data <- data[-i, ]
    removed <- urban.pcadata[i, ncol(urban.pcadata)][[1]]
    usesites <- sites[-i]
    removed_site <- sites[i]
  }
  result <- run_pls(data)
  loadings <- result[["loadings"]]
  plsdata <- result[["plsdata"]]
  coef <- result[["coef"]]
  landuse_type <- names(loadings)

  resloadings <- tibble(landuse_type, loadings)
  resloadings$p_removed <- removed
  resloadings$site_removed <- removed_site
  jkloadings[[paste(i)]] <- resloadings

  resplsdata <- plsdata
  resplsdata$p_removed <- removed
  resplsdata$site <- usesites
  jkplsdata[[paste(i)]] <- resplsdata
  coefres <- tibble(coef, removed, usesites)
  jkcoefs[[paste(i)]] <- coefres
}

# process results of jackknife for further analysis
urban_jk_loadings <- do.call(bind_rows, jkloadings)
urban_jk_loadings$cutoff <- sqrt(
  1 / (length(unique(urban_jk_loadings$landuse_type)) - 1)
)
urban_jk_pls <- do.call(bind_rows, jkplsdata)
urban_jk_coef <- do.call(bind_rows, jkcoefs)
```



### Identifying outlier sites from the slopes of the jackknifed beta regression

```{r, urban_jk_slope_dist, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
# show coef calculated with peticular prop scouts removed from analysis
urban_jk_coef <- do.call(bind_rows, jkcoefs)
hist(urban_jk_coef$coef, col = "pink")
```

```{r, urban_jk_slope_dist_value, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
urban_jk_coef %>% filter(coef < 0.4)
```


```{r, urban_jk_outlier_removed, message = FALSE, error = FALSE, warning = FALSE, echo = TRUE}

# full data fit
model_pls <- plsRbeta(
  p_transform ~ .,
  data = urban.pcadata,
  nt = ncol(urban.pcadata) - 1,
  model = "pls-beta",
  verbose = FALSE
)

print(model_pls)

plsdata <- data.frame(
  prop.scouts = model_pls$dataY,
  pc1 = model_pls$tt[, 1]
)
plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  plsdata
)

print("Full data beta regression")
print(summary(plsbetamodel))

# set up plot of pls fit
newdat <- data_frame(
  pc1 = seq(
    min(plsdata$pc1),
    max(plsdata$pc1),
    length.out = length(plsdata$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
plsdata$p <- (2 * n * plsdata$prop.scouts - 1) / (2 * (n - 1))

# make plot of pls fit
plt_full <- ggplot(data = plsdata, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  )

# plot loadings
full_loadings <- as_tibble(model_pls$pp[, 1])
full_loadings$landuse_type <- names(model_pls$pp[, 1])

urban_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
full_loading_plt <- ggplot(data = full_loadings, aes(
  x = landuse_type, y = value, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = value, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#BBBBBB", other = "#BBBBBB"
    )
  )

# outlier removed fit
removed_data <- urban.pcadata %>% filter(p_transform < 0.95)

model_pls <- plsRbeta(
  p_transform ~ .,
  data = removed_data,
  nt = ncol(removed_data) - 1,
  model = "pls-beta",
  verbose = FALSE
)
plsdata <- data.frame(
  prop.scouts = model_pls$dataY,
  pc1 = model_pls$tt[, 1]
)
plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  plsdata
)

print("Outlier removed beta regression")
print(summary(plsbetamodel))

# set up plot of pls fit
newdat <- data_frame(
  pc1 = seq(
    min(plsdata$pc1),
    max(plsdata$pc1),
    length.out = length(plsdata$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
plsdata$p <- (2 * n * plsdata$prop.scouts - 1) / (2 * (n - 1))

# make plot of pls fit
plt_outlier_removed <- ggplot(data = plsdata, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  ) +
  scale_y_continuous(limits = c(0, 1))

# plot loadings
outlier_rmv_loadings <- as_tibble(model_pls$pp[, 1])
outlier_rmv_loadings$landuse_type <- names(model_pls$pp[, 1])

outlier_rmv_loading_plt <- ggplot(data = outlier_rmv_loadings, aes(
  x = landuse_type, y = value, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = value, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_colour_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )


# show all plots
plot_grid(
  plt_full, plt_outlier_removed,
  full_loading_plt, outlier_rmv_loading_plt,
  ncol = 2, labels = "AUTO"
)
```

### Examining how the loadings and fit changes through the jackknifed analysis

```{r, urban_jk_averaged_fits, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

# plot loadings
urban_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
urban_average_loadings_plt <- ggplot(data = urban_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "continuous.central" = "#BBBBBB", "dense.residential" = "#BBBBBB",
      "sparse.residential" = "#EE6677", "railway" = "#CCBB44",
      "parks.allotments.cemeteries" = "#228833",
      "water" = "#BBBBBB", other = "#BBBBBB"
    )
  )


# process plsdata for model fit on the median or mean response from JK PLS
urban_jk_pls <- do.call(bind_rows, jkplsdata)
n <- 10
urban_jk_pls$p <- (2 * n * urban_jk_pls$prop.scouts - 1) / (2 * (n - 1)) # convert to p

forjoin <- urban_lu %>%
  select(site, p)

urban_median_pls_data <- urban_jk_pls %>%
  group_by(site) %>%
  summarise(pc1 = median(Comp_.1, na.rm = TRUE)) %>%
  right_join(forjoin, by = "site") %>%
  mutate(prop.scouts = (p * (n - 1) + 0.5) / n)

plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  urban_median_pls_data
)

# show pls summary
print(summary(plsbetamodel))

# set up plot of pls average fit
newdat <- data_frame(
  pc1 = seq(
    min(urban_median_pls_data$pc1),
    max(urban_median_pls_data$pc1),
    length.out = length(urban_median_pls_data$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
urban_median_pls_data$p <- (2 * n * urban_median_pls_data$prop.scouts - 1) / (2 * (n - 1))

# make plot of average pls fit
urban_average_fit_plt <- ggplot(data = urban_median_pls_data, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  )

# show all plots
plot_grid(
  urban_average_fit_plt, urban_average_loadings_plt,
  ncol = 1, labels = "AUTO"
)
```

## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings

```{r, urban_jk_loadings_all, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

make_plots_func <- function(urban_jk_loadings) {
  plt <- ggplot(data = urban_jk_loadings, aes(
    x = landuse_type, y = loadings, fill = landuse_type
  )) +
    geom_point(aes(colour = landuse_type), size = 8) +
    geom_segment(aes(
      x = landuse_type, xend = landuse_type,
      y = 0, yend = loadings, colour = landuse_type, size = .1
    )) +
    geom_hline(
      yintercept = 0, color = "black"
    ) +
    annotate(
      "rect",
      ymin = -urban_cuttoff,
      ymax = urban_cuttoff,
      xmin = 0, xmax = Inf,
      alpha = 0.2
    ) +
    ylab("Correlation") +
    coord_flip() +
    theme_classic() +
    theme(
      axis.line.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = "None",
      axis.title.x = element_text(size = 80),
      axis.text.x = element_text(size = 60)
    ) +
    scale_colour_manual(
      values = c(
        "sparse.residential" = "#EE6677",
        "railway" = "#CCBB44", "parks.allotments.cemeteries" = "#228833"
      )
    ) +
    facet_wrap(~site_removed, nrow = 1) +
    theme(
      strip.text = element_text(size = 80)
    )
  return(plt)
}

plt <- make_plots_func(urban_jk_loadings %>% filter(site_removed != "None"))

ggsave(
  plot = plt,
  filename = "../Results/figures/SI_figures/urban_jk_individual_loadings.png",
  width = 800,
  height = 400,
  units = "mm",
  dpi = 300
)

# bur_blo <- urban_jk_loadings %>% filter(site_removed %in% c("BLO", "BUR")) %>%
#   make_plots_func()

# ggsave(
#   plot = bur_blo,
#   filename = "../Results/figures/SI_figures/bur_blo_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# cad_gil <- urban_jk_loadings %>% filter(site_removed %in% c("CAD", "GIL")) %>%
#   make_plots_func()

# ggsave(
#   plot = cad_gil,
#   filename = "../Results/figures/SI_figures/cad_gil_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# hhs_hor <- urban_jk_loadings %>% filter(site_removed %in% c("HHS", "HOR")) %>%
#   make_plots_func()

# ggsave(
#   plot = hhs_hor,
#   filename = "../Results/figures/SI_figures/hhs_hor_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# mak_mpa <- urban_jk_loadings %>% filter(site_removed %in% c("MAK", "MPA")) %>%
#   make_plots_func()

# ggsave(
#   plot = mak_mpa,
#   filename = "../Results/figures/SI_figures/mak_mpa_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# swp_zsl <- urban_jk_loadings %>% filter(site_removed %in% c("SWP", "ZSL")) %>%
#   make_plots_func()

# ggsave(
#   plot = swp_zsl,
#   filename = "../Results/figures/SI_figures/swp_zsl_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

```

## How does this compare with the proportion of land-coverage for each site?

```{r, urban_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

area_sites_used <- urban_lu %>% filter(site != "CAD")

total_area <- area_sites_used[, 3:ncol(area_sites_used)] %>%
  colSums()
total_area <- tibble(landuse_type = names(total_area), area = total_area)
total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100

urban_lu_loadings_with_area <- urban_jk_loadings %>%
  left_join(total_area, by = "landuse_type") %>%
  filter(site_removed == "CAD") # <---- remove outlier site

ggplot(data = urban_lu_loadings_with_area, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type, size = area_pct)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = loadings, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )

urban_lu_loadings_with_area %>%
  arrange(-area_pct) %>%
  print()
```

```{r, urban_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

fit1 <- lm(area ~ loadings, data = urban_lu_loadings_with_area)
print(summary(fit1))

print(
  ggplotRegression(fit1)
)
```

### Conclusions

Site CAD seems to be having a large effect on the results of the PLS. The jackknife analysis shows this site skews the slopes more than the others. Plotting the averages of the PLS provides a means of seeing how the PLS results change in relation to removing sites. They seem fairly robust to site removals, but this could also mean site CAD is heavily influencing the relationship.


## Agri-rural PLS


```{r, rural_pls, message = FALSE, error = FALSE, warning = FALSE, include = FALSE}

# set up data for pls
rural.pcadata <- rural_lu[, 3:length(names(rural_lu))]
sites <- rural_lu$site
rural_cuttoff <- sqrt(1 / (ncol(rural.pcadata) - 1))
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n
rural.pcadata$p_transform <- y

# create empty lists to store jackknifed results in
jkloadings <- list()
jkplsdata <- list()
jkcoefs <- list()

# jackknife the pls analysis
for (i in c(0, seq_len(nrow(rural.pcadata)))) {
  data <- rural.pcadata
  removed <- 0
  usesites <- "None"
  removed_site <- "None"
  if (i > 0) {
    data <- data[-i, ]
    removed <- rural.pcadata[i, ncol(rural.pcadata)][[1]]
    usesites <- sites[-i]
    removed_site <- sites[i]
  }
  result <- run_pls(data)
  loadings <- result[["loadings"]]
  plsdata <- result[["plsdata"]]
  coef <- result[["coef"]]
  landuse_type <- names(loadings)

  resloadings <- tibble(landuse_type, loadings)
  resloadings$p_removed <- removed
  resloadings$site_removed <- removed_site
  jkloadings[[paste(i)]] <- resloadings

  resplsdata <- plsdata
  resplsdata$p_removed <- removed
  resplsdata$site <- usesites
  jkplsdata[[paste(i)]] <- resplsdata
  coefres <- tibble(coef, removed, usesites)
  jkcoefs[[paste(i)]] <- coefres
}

# process results of jackknife for further analysis
rural_jk_loadings <- do.call(bind_rows, jkloadings)
rural_jk_loadings$cutoff <- sqrt(
  1 / (length(unique(rural_jk_loadings$landuse_type)) - 1)
)
save(urban_jk_loadings, rural_jk_loadings, file = "../Results/pls-loadings.RData")
rural_jk_pls <- do.call(bind_rows, jkplsdata)
rural_jk_coef <- do.call(bind_rows, jkcoefs)
```


```{r, rural_jk_slope_dist, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}
# show coef calculated with peticular prop scouts removed from analysis
rural_jk_coef <- do.call(bind_rows, jkcoefs)
jkci <- quantile(rural_jk_coef$coef, probs = c(0.025, .975))
hist(rural_jk_coef$coef, col = "pink")
```

### Examining how the loadings and fit changes through the jackknifed analysis

```{r, rural_jk_averaged_fits, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

# plot loadings
rural_loading_text <- tibble(landuse_type = unique(landuse_type), result = 0)
rural_average_loadings_plt <- ggplot(data = rural_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_boxplot() +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -rural_cuttoff,
    ymax = rural_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )


# process plsdata for model fit on the median or mean response from JK PLS
rural_jk_pls <- do.call(bind_rows, jkplsdata)
n <- 10
# convert back to p
rural_jk_pls$p <- (2 * n * rural_jk_pls$prop.scouts - 1) / (2 * (n - 1))

forjoin <- rural_lu %>%
  select(site, p)

rural_median_pls_data <- rural_jk_pls %>%
  group_by(site) %>%
  summarise(pc1 = median(Comp_.1, na.rm = TRUE)) %>%
  right_join(forjoin, by = "site") %>%
  mutate(prop.scouts = (p * (n - 1) + 0.5) / n)

plsbetamodel <- betareg(
  prop.scouts ~ pc1,
  rural_median_pls_data
)

# show pls summary
print(summary(plsbetamodel))

# set up plot of pls average fit
newdat <- data_frame(
  pc1 = seq(
    min(rural_median_pls_data$pc1),
    max(rural_median_pls_data$pc1),
    length.out = length(rural_median_pls_data$pc1)
  )
)
newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# get effects
fitdata <- as.data.frame(predictorEffects(plsbetamodel))$pc1
rural_median_pls_data$p <- (2 * n * rural_median_pls_data$prop.scouts - 1) / (2 * (n - 1))

# make plot of average pls fit
rural_average_fit_plt <- ggplot(data = rural_median_pls_data, aes(x = pc1, y = p)) +
  geom_point() +
  geom_line(data = fitdata, aes(x = pc1, y = fit)) +
  geom_ribbon(
    data = fitdata,
    aes(x = pc1, y = fit, ymin = lower, ymax = upper),
    alpha = 0.2,
    fill = "deepskyblue"
  )


# show all plots
plot_grid(
  rural_average_fit_plt, rural_average_loadings_plt,
  ncol = 1, labels = "AUTO"
)
```


## Further analysis of sensitivity to site removal and landuse proportions: digging into the the distributions of PLS loadings

```{r, rural_jk_loadings_all, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

make_plots_func2 <- function(rural_jk_loadings){
  plt <- ggplot(data = rural_jk_loadings, aes(
    x = landuse_type, y = loadings, fill = landuse_type
  )) +
    geom_point(aes(colour = landuse_type), size = 8) +
    geom_segment(aes(
      x = landuse_type, xend = landuse_type,
      y = 0, yend = loadings, colour = landuse_type, size = .1
    )) +
    geom_hline(
      yintercept = 0, color = "black"
    ) +
    annotate(
      "rect",
      ymin = -rural_cuttoff,
      ymax = rural_cuttoff,
      xmin = 0, xmax = Inf,
      alpha = 0.2
    ) +
    ylab("Correlation") +
    coord_flip() +
    theme_classic() +
    theme(
      axis.line.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks.y = element_blank(),
      legend.position = "None",
      axis.title.x = element_text(size = 80),
      axis.text.x = element_text(size = 60)
    ) +
    scale_colour_manual(
      values = c(
        "arable" = "#228833", "built.up.area" = "#EE6677",
        "Non.agricultural.unmanaged.green.space" = "#CCBB44",
        "water" = "#4477AA"
      )
    ) +
    facet_wrap(~site_removed, nrow = 1) +
    theme(
      strip.text = element_text(size = 80)
    ) +
    xlab("PC1 Loadings")
  return(plt)
}

plt <- make_plots_func2(rural_jk_loadings %>% filter(site_removed != "None"))

ggsave(
  plot = plt,
  filename = "../Results/figures/SI_figures/rural_jk_individual_loadings.png",
  width = 800,
  height = 400,
  units = "mm",
  dpi = 300
)

# bel_bfi <- rural_jk_loadings %>% filter(site_removed %in% c("BEL", "BFI")) %>%
#   make_plots_func2()

# ggsave(
#   plot = bel_bfi,
#   filename = "../Results/figures/SI_figures/bel_bfi_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# her_mel <- rural_jk_loadings %>% filter(site_removed %in% c("HER", "MEL")) %>%
#   make_plots_func2()

# ggsave(
#   plot = her_mel,
#   filename = "../Results/figures/SI_figures/her_mel_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# sau_som <- rural_jk_loadings %>% filter(site_removed %in% c("SAU", "SOM")) %>%
#   make_plots_func2()

# ggsave(
#   plot = sau_som,
#   filename = "../Results/figures/SI_figures/sau_som_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )

# sra_stu_yal <- rural_jk_loadings %>% filter(site_removed %in% c("SRA", "STU", "YAL")) %>%
#   make_plots_func2()

# ggsave(
#   plot = sra_stu_yal,
#   filename = "../Results/figures/SI_figures/sra_stu_yal_loadings.png",
#   width = 120,
#   height = 200,
#   units = "mm",
#   dpi = 300
# )


```

## How does this compare with the proportion of land-coverage for each site?

```{r, rural_jk_loadings_landuse_area, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

total_area <- rural_lu[, 3:ncol(rural_lu)] %>%
  colSums()
total_area <- tibble(landuse_type = names(total_area), area = total_area)
total_area$area_pct <- (total_area$area / sum(total_area$area)) * 100

rural_lu_loadings_with_area <- rural_jk_loadings %>%
  left_join(total_area, by = "landuse_type") %>%
  filter(site_removed == "None")

ggplot(data = rural_lu_loadings_with_area, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type, size = area_pct)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = loadings, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -rural_cuttoff,
    ymax = rural_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_colour_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  )

rural_lu_loadings_with_area %>%
  arrange(-area_pct) %>%
  print()
```


```{r, rural_jk_loadings_landuse_area_correlation, message = FALSE, error = FALSE, warning = FALSE, echo = FALSE}

fit1 <- lm(area ~ loadings, data = rural_lu_loadings_with_area)
print(summary(fit1))

print(
  ggplotRegression(fit1)
)
```
